---
title: "Club convergence of regional price in 34 Indonesian provinces"
author: "Harry Aginta"
output:
  html_notebook:
    code_folding: show
    highlight: monochrome
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document: default
  html_document:
    code_download: true
    df_print: paged
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
    number_sections: true
    code_folding: "show"
    theme: "cosmo"
    highlight: "monochrome"
  word_document: default
---

```{r}
#install.packages("ConvergenceClubs")
#install.packages("mFilter")
#install.packages("data.table")
#install.packages("plyr")
#install.packages("tidyverse")
#install.packages("tibble")
#install.packages("janitor")
#install.packages("oglmx") # For ordered logit model in R
#install.packages("rms") # # For ordered logit model in R too as an alternative
library(ConvergenceClubs)
library(mFilter)
library(tidyverse)
library(tibble)
library(data.table)
library(plyr)
library(dplyr)
library(janitor)
library(oglmx)
library(rms)
```

### Load the data
```{r}
cpi <- rio::import("https://github.com/quarcs-lab/Tutorial-club-convergence-regional-price-Indonesia_Harry-Ch2/blob/main/data/CPI_province.xlsx?raw=true")
cpi
```

```{r}
glimpse(cpi)
```

### Remove ID column from data
```{r}
cpi <- subset(cpi, select = -1)
```


# Part A: Visualize CPI all provinces
## Reshape to long df 
```{r}
cpi_long <- cpi %>% 
  pivot_longer(-Province, names_to = "date", values_to = "cpi") 
cpi_long
```
## Convert numeric date from Excel to date R
```{r}
cpi_long$date <- excel_numeric_to_date(as.numeric(as.character(cpi_long$date)), date_system = "modern")
cpi_long
```


## Boxplot by Province
```{r}
boxplot.cpi <- ggplot(cpi_long, aes(x = Province, y = cpi))
boxplot.cpi + geom_boxplot() + ggtitle("CPI by Province") +
  xlab("") + ylab("CPI") + theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) + theme(axis.text.x = element_text(size = rel(1), angle = 90, hjust = 1))
#ggsave("boxplot_cpi.png") #Save the boxplot in .png
```


## Plot CPIs using ggplot
```{r}
graph <- cpi_long %>%
ggplot(aes(x = date, y = cpi, color=Province)) +
geom_line(size=1) +
labs(subtitle = "Consumer Price Index",
x = "Time",
y = "CPI", color="") +
theme(text=element_text( family="Palatino")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  guides(col = guide_legend(nrow = 8))
graph
#ggsave("graph_cpi.png") #Save the graph in .png
```


# Part B: Club convergence
## Reshape data from long to wide. The club convergence package requires datafarame in wide format.
```{r}
cpi_wide <- cpi_long %>%
  pivot_wider(names_from = date, values_from = cpi)
cpi_wide
```

## Transform CPI to log and then remove short-run fluctuation using HP filter
```{r}
logcpi <- log(cpi_wide[,-1])
```


```{r}
filteredcpi <- apply(logcpi, 1, 
function(x){mFilter::hpfilter(x, freq=14400, type="lambda")$trend} ) 
filteredcpi <- data.frame(Province = cpi[,1], t(filteredcpi), stringsAsFactors=FALSE ) 
colnames(filteredcpi) <- colnames(cpi_wide) 
filteredcpi
```

## Apply the log-t-test
```{r}
H <- computeH(filteredcpi[,-1], quantity = "H")
round(estimateMod(H, time_trim=0.333, HACmethod = "FQSB"), 3)
```

Ho: There is overall convergence. We can reject Ho given beta coefficient is negative and the tvalue is smaller than -1.65.
Then we proceed to identify convergence clubs.


## Identify clubs
```{r}
clubs <- findClubs(filteredcpi, dataCols=2:97, unit_names = 1, refCol=97,
time_trim=0.333, cstar=0, HACmethod = 'FQSB')
```

```{r}
summary(clubs)
```

```{r}
print(clubs)
```

```{r}
plot(clubs)
```

```{r}
plot(clubs, clubs=NULL, avgTP = TRUE, legend=TRUE)
```
NOTE: We observe 5 club convergence.


## Merge clubs
```{r}
mclubs <- mergeClubs(clubs, mergeMethod='PS')
```

```{r}
summary(mclubs)
```

```{r}
mclubs
```

```{r}
plot(mclubs)
```

  
```{r}
plot(mclubs, clubs=NULL, avgTP = TRUE, legend=TRUE)
```
NOTE: We observe 4 final club convergence, where club 2 and 3 merge.


# Part C: Visualize the transition path of club convergence
## Convert the "mclubs" object from list to dataframe.

```{r}
class(mclubs)
```

```{r}
table.mclubs <- map(mclubs, as.data.table)
df.mclubs <- rbindlist(table.mclubs, fill = T, idcol = T)
df.mclubs
```

## Change the the name of the 1st and 5th column from to "Clubs" and "Province"
```{r}
colnames(df.mclubs)[c(1,5)] <- c("Club","Province")
df.mclubs
```

## Merge "df.mclubs" with "filteredcpi" to get the filtered CPI data of all provinces 
```{r}
cpi.club <- as.data.frame(
  inner_join(filteredcpi, df.mclubs, by="Province") %>% 
  select(-contains(c("clubs","id","model")))
  ) 
cpi.club
```

## Generate the list of relative CPI
```{r}
rel.cpi <- list()
for(a in 2:97) {
  rel.cpi[[a]] <- data.frame(cpi.club$Province, cpi.club[,a]/mean(cpi.club[,a]))
colnames(rel.cpi[[a]])[1]<- "Province"
colnames(rel.cpi[[a]])[2]<- 
paste("rel",colnames(cpi.club)[a],sep="_")
}

for (x in 2:96) {
  rel.cpi[[x+1]]<-left_join(rel.cpi[[x]],rel.cpi[[x+1]], by="Province")
}
```

## Convert the list to dataframe
```{r}
table.rel.cpi <- map(rel.cpi, as.data.table)
df.rel.cpi <- rbindlist(table.rel.cpi, fill = T, idcol = T) %>% drop_na()

df.rel.cpi <- join(df.rel.cpi,cpi.club,by=c('Province')) %>% 
  select(Province,contains("rel"),Club)
  
df.rel.cpi <- as.data.frame(df.rel.cpi)
df.rel.cpi
```

## Remove "rel_" string from colums' name to retain only date
```{r}
colnames(df.rel.cpi) <- gsub("rel_","",colnames(df.rel.cpi))
df.rel.cpi
```

## Transpose to long format data for ggplot
```{r}
df.rel.cpi.long <- df.rel.cpi %>%  pivot_longer(-c(Province,Club), names_to = "Time", values_to="Rel_CPI")
df.rel.cpi.long
```

## Compute the relative values by clubs for each month and remove observation in 2012 to exclude index base year in the plot.
```{r}
df.rel.cpi.path <- aggregate(Rel_CPI ~ Club + Time, df.rel.cpi.long, mean) 
df.rel.cpi.path <- df.rel.cpi.path[-c(1:48), ]
df.rel.cpi.path$Time <- as.Date(as.character(df.rel.cpi.path$Time))
df.rel.cpi.path
```
```{r}
path_all <- df.rel.cpi.path %>% 
  ggplot(aes(x=Time,y=Rel_CPI, group=Club, col=Club)) + geom_line() +
  labs(title = "Transition path of all clubs") +
  theme_bw()
path_all
#ggsave("path_all.png") #save the graph in .png
```

# Part D: Investigate the influcencing factors of club convergence using ordered logit model
## Load the data
```{r}
library(readr)
ologit <- read_delim("https://raw.githubusercontent.com/quarcs-lab/Tutorial-club-convergence-regional-price-Indonesia_Harry-Ch2/main/data/ologit.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)
ologit
```

## Generate new variable of log(labor productivity)
```{r}
ologit <- ologit %>% 
mutate("lnlbr_prod" = log(lbr_prod))
ologit
```

## Order the categorical club to get "the higher club_status, the higher CPI club".
```{r}
ologit$club_status <- 4
ologit$club_status[ologit$ID_Club == 2] <- 3
ologit$club_status[ologit$ID_Club == 3] <- 2
ologit$club_status[ologit$ID_Club == 4] <- 1
ologit
```

## Define Y and X variables
```{r}
attach(ologit)
Y <- cbind(club_status)
X <- cbind(lnlbr_prod, backinf, spainfquepoly19, pcons_g, gexp_g)
Xvar <- c("lnlbr_prod", "backinf", "spainfquepoly19", "pcons_g", "gexp_g")
```
NOTE: 
"lnlbr_prod" = log of labor productivity
"backinf" = backward inflation expectation
"spainfquepoly19" = spatial lag of inflation
"pcons_g" = growth of personal consumption
"gexp_g" = growth of government expenditure. 


## Run ordered logit model using ologit.reg command in oglmx package
```{r}
model.logit <- ologit.reg(Y ~ X, data = ologit)
summary(model.logit)
```

## Run ordered logit model using lrm command in rms package
```{r}
ddist <- datadist(Xvar)
options(datadist="ddist")
```

```{r}
m <- lrm(Y ~ X, data = ologit)
print(m)
```
